Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-29 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Revised Chapter Outline

\chapter{Scope of Study}
\label{chapter:Scope of Study}

@@

\chapter{Technical Approach}
\label{chapter:Technical Approach}

@@

\section{Selection of Measurement Methods and Analytical Tools}

@@

\section{Fire Pattern Experiments}

@@

\section{Bench Scale Experiments}

@@

@@

\section{Fire Pattern Recreation}

@@

\chapter{Results}
\label{chapter:Results}

{\section{HRR}

@@

\section{Plume Temperatures}

@@

\section{Total Heat Flux}

@@

\section{Mass Loss}

@@

\section{Flame Height}

@@

\section{Summary}

@@

\chapter{Instrumented Calibration (non-combustible) Wall Experiments}

@@

Two vertical arrays of water cooled, Schmidt-Boelter total heat flux gauges, as previously described in the instrumentation section above, were installed at 20 cm intervals from 0.2~m to 1.2~m above the burner.  One vertical array was centered on the burner and the other vertical array was centered on the right edge of the burner. The arrangement is shown in Fig.~\ref{Instrumented_NC_Wall}. Bare-bead, Chromel-Alumel (type K) thermocouples, with a 0.5~mm nominal diameter were installed next to the center of each heat flux sensors and continued, at 20~cm intervals, up to a height of 2.2~m above the surface of the burner. Photographs of the front and rear of the instrumented wall are shown in Figures~\ref{Instrumented_Wall_Front_photo} and \ref{Instrumented_Wall_Rear_photo} The thermocouples where installed in small holes from the back of the wall of the wall.  The bead of the thermocouple was flush with the surface on the calcium silcate board and was in contact with the board, as shown in Figure~\ref{Instrumented_Wall_Close_up_TC_HF_photo} For the center array,a second thermocouple was installed on the backside of the calcium silicate board, approximately 10~mm below the TC positioned on the front surface of the wall. The photograph in Figure~\ref{Instrumented_Wall_Detail_Rear_TC_photo} shows a rear thermocouple with the bead in contact with the rear surface of the wall and the relative position to the thermocouple on the front surface of the wall. In a manner similar to the source fire characterization experiments, the instrumented wall experiments were conducted under the NIST Large Fire Research Laboratory’s 3~m by 3~m oxygen depletion calorimeter.

@@

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix X.  For the natural gas fueled experiments only two replicates were conducted at each position given the higher level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below, examples of the averaged data from the gasoline fueled experiments will be presented to demonstrate the impact on distance on the conditions measured on the wall. Then a composite graph of the measured thermal impacts on the wall versus distance will be shown for each fuel type.

@@

Figure~\ref{IW_SC_TWGAS} shows the temperatures of the array of thermocouples located along the vertical centerline of the gasoline fueled burner.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average temperature of the replicate experiments, for each thermocouple position is shown.  Moving from 0.2~m above the top surface of the burner up to 2.2~m above the top of the burner surface in increments of 0.2~m.

\begin{figure}[p]
	\centering
	\includegraphics[width=2.5in]{../Figures/TC_Surface_Center_TWGas_0_61m_RepsAvg}
	\includegraphics[width=2.5in]{../Figures/TC_Surface_Center_TWGas_0_31m_RepsAvg}\\
	\includegraphics[width=2.5in]{../Figures/TC_Surface_Center_TWGas_0_15m_RepsAvg}
	\includegraphics[width=2.5in]{../Figures/TC_Surface_Center_TWGas_0m_RepsAvg}\\
	\caption[Wall Temperatures along the centerline of the gasoline fueled burner]{Wall Temperatures along the centerline of the gasoline fueled burner. The temperatures are averages of three replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top left position and moving from left to right the distances are 0.61~m, 0.31~m,0.15~m,and 0~m.}
	\label{IW_SC_TWGAS}
\end{figure}

@@

Figure~\ref{IW_HFC_TWGAS} shows the temperatures of the array of total heat flux gauges located along the vertical centerline of the gasoline fueled burner.  The sensing surface of the gauges were installed flush with the wall surface and exposed to the fire. In each graph, the average temperature of the replicate experiments, for each total heat flux position is shown.  Moving from 0.2~m above the top surface of the burner up to 1.2~m above the top of the burner surface in increments of 0.2~m.

\begin{figure}[p]
	\centering
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_61m_Reps_Avg}
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_31m_Reps_Avg}\\
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_15m_Reps_Avg}
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0m_Reps_Avg}\\
	\caption[Wall heat fluxes along the centerline of the gasoline fueled burner]{Wall heat fluxes along the centerline of the gasoline fueled burner.  The heat fluxes are averages of three replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top left position and moving from left to right the distances are 0.61~m, 0.31~m,0.15~m,and 0~m.}
	\label{IW_HFC_TWGAS}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/NCTW_NG_HF_Center_Avg}\\
	\caption[Summary of wall heat fluxes along the centerline of the natural gas fueled burner]{Summary of the heat flux to the non-combustible wall along the verticle centerline of the natural gas fueled burner compared to the position away from the wall.}
	\label{nctw_ng_heatflux}
\end{figure}

>>>>

\section{Scope of Study}

@@

\section{Technical Approach}

@@

\subsection{Selection of Measurement Methods and Analytical Tools}

@@

\subsection{Fire Pattern Experiments}

@@

\subsection{Bench Scale Experiments}

@@

\subsubsection{Critical Heat Flux}

@@

\subsection{Numerical Fire Pattern Analysis}

@@

\subsection{HRR}

@@

@@

\subsection{Total Heat Flux}

@@

\subsection{Mass Loss}

@@

@@

\subsection{Summary}

@@

\chapter{Instrumented Non-combustible Wall Experiments}

@@

Two vertical arrays of water cooled, Schmidt-Boelter total heat flux gauges, as previously described in the instrumentation section above, were installed at 20 cm intervals from 0.2~m to 1.2~m above the burner.  One vertical array was centered on the burner and the other vertical array was centered on the right edge of the burner. The arrangement is shown in Fig.~\ref{Instrumented_NC_Wall}. Bare-bead, Chromel-Alumel (type K) thermocouples, with a 0.5~mm nominal diameter were installed next to the center of each heat flux sensors and continued, at 20~cm intervals, up to a height of 2.2~m above the surface of the burner. Photographs of the front and rear of the instrumented wall are shown in Figures~\ref{Instrumented_Wall_Front_photo} and \ref{Instrumented_Wall_Rear_photo} The thermocouples where installed in small holes from the back of the wall of the wall.  The bead of the thermocouple was flush with the surface on the calcium silcate board and was in contact with the board, as shown in Figure~\ref{Instrumented_Wall_Close_up_TC_HF_photo} For the center array,a second thermocouple was installed on the backside of the calcium silicate board, approximately 10~mm below the TC positioned on the front surface of the wall. The photograph in Figure~\ref{Instrumented_Wall_Detail_Rear_TC_photo} shows a rear thermocouple with the bead in contact with the rear surface of the wall and the relative position to the thermocouple on the front surface of the wall. In a manner similar to the source fire characterization experiments, the instrumented wall experiments were conducted under the NIST Large Fire Research Laboratory's 3~m by 3~m oxygen depletion calorimeter.

@@

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix X.  For the natural gas fueled experiments only two replicates were conducted at each position given the higher level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below,a full set of the averaged data from the natural gas fueled experiments will be presented to demonstrate the impact on distance from the wall on the temperatures in the plume as well as the conditions measured on the wall. Then composite graphs of the wall temperatures and heat flux measure on the wall versus distance will be shown for each fuel type.

Figures~\ref{TC_Plume_TWNG_comp} through \ref{TC_Back_Center_TWNG_comp} show a complete set of data, averaged over two replicates, for the instrumented, non combustible wall natural gas burner experiments.

\begin{figure}[p]
	\centering
	\includegraphics[width=3.in]{../Figures/TC_Plume_TWNG_0_61m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Plume_TWNG_0_31m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Plume_TWNG_0_15m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Plume_TWNG_0m_Reps_Avg}\\
	\caption[Plume temperatures along the centerline of the natural gas fueled burner]{Plume temperatures along the centerline of the natural gas fueled burner. The temperatures are averages of two replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down the distances from the wall are 0.61~m, 0.31~m, 0.15~m, and 0~m respectively.}
	\label{TC_Plume_TWNG_comp}
\end{figure}

\begin{figure}[p]
	\centering
		\includegraphics[width=3.in]{../Figures/TC_Surface_Center_TWNG_0_61m_Reps_Avg}\\
		\includegraphics[width=3.in]{../Figures/TC_Surface_Center_TWNG_0_31m_Reps_Avg}\\
		\includegraphics[width=3.in]{../Figures/TC_Surface_Center_TWNG_0_15m_Reps_Avg}\\
		\includegraphics[width=3.in]{../Figures/TC_Surface_Center_TWNG_0m_Reps_Avg}\\
	\caption[Wall Temperatures along the centerline of the natural gas fueled burner]{Wall Temperatures along the centerline of the natural gas fueled burner. The temperatures are averages of two replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down the distances from the wall are 0.61~m, 0.31~m, 0.15~m, and 0~m, respectively.}
	\label{TC_Surf_Cent_TWNG_comp}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.in]{../Figures/TC_Surface_Offset_TWNG_0_61m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Surface_Offset_TWNG_0_31m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Surface_Offset_TWNG_0_15m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Surface_Offset_TWNG_0m_Reps_Avg}\\
	\caption[Wall Temperatures along the edge of the natural gas fueled burner]{Wall Temperatures along the right edge of the natural gas fueled burner, when facing the wall. The temperatures are averages of two replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down the distances from the wall are 0.61~m, 0.31~m, 0.15~m, and 0~m, respectively.}
	\label{TC_Surf_Edge_TWNG_comp}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.in]{../Figures/TC_Back_Center_TWNG_0_61m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Back_Center_TWNG_0_31m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Back_Center_TWNG_0_15m_Reps_Avg}\\
	\includegraphics[width=3.in]{../Figures/TC_Back_Center_TWNG_0m_Reps_Avg}\\
	\caption[Backside wall Temperatures along the centerline of the natural gas fueled burner]{Backside wall Temperatures along the centerline of the natural gas fueled burner. The temperatures are averages of two replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down the distances from the wall are 0.61~m, 0.31~m, 0.15~m, and 0~m, respectively.}
	\label{TC_Back_Center_TWNG_comp}
\end{figure}

@@

Figure~\ref{NCTW_Surf_Center_comp} shows the temperatures of the array of thermocouples located along the vertical centerline of the burners for each of the three fuels.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average peak temperature of the replicate experiments, for each thermocouple position is shown. Moving from 0.2~m above the top surface of the burner up to 2.2~m above the top of the burner surface in increments of 0.2~m. As in the data analysis for the source fires, the peak averages were based on a 30 s period bounding the peak for the natural gas and gasoline fires.  A 15 s interval bounding the peak was used for the polyurethane, in order to be more representative of the peak conditions.

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/NCTW_NG_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_GAS_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_PUF_TC_Surface_Center_Avg}\\

	\caption[Composite graph of the wall temperatures along the centerline of the burners]{Composite graph of the wall temperatures along the centerline of the burners for each fuel. The peak temperatures are averages of replicate experiments for each instrument location at each position from the wall. Beginning with the top graph position and moving down the graph for natural gas, gasoline, and polyurethane foam are presented.}
	\label{NCTW_Surf_Center_comp}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/NCTW_NG_TC_Surface_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_GAS_TC_Surface_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_PUF_TC_Surface_Offset_Avg}\\

	\caption[Composite graph of the wall temperatures along the edge of the burners]{Composite graph of the wall temperatures along the edge of the burners for each fuel. The peak temperatures are averages of replicate experiments for each instrument location at each position from the wall. Beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam are presented.}
	\label{NCTW_Surf_Edge_comp}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/NCTW_NG_TC_Back_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_GAS_TC_Back_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_PUF_TC_Back_Center_Avg}\\

	\caption[Composite graph of the backside wall temperatures along the center of the burners]{Composite graph of the backside wall temperatures along the vertical centerline of the burners for each fuel. The peak temperatures are averages of replicate experiments for each instrument location at each position from the wall. Beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam are presented.}
	\label{NCTW_Back_Cent_comp}
\end{figure}

@@

Figure~\ref{NCTW_HF_Center_Comp} shows the results from the array of total heat flux gauges located along the vertical centerline of the burners for each fuel.  The sensing surface of the gauges were installed flush with the wall surface and exposed to the fire. In each graph, the averaged peak heat flux for each of the replicate experiments and each position is shown.  Moving from 0.2~m above the top surface of the burner up to 1.2~m above the top of the burner surface in increments of 0.2~m.

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/NCTW_NG_HF_Center_Avg} \\
	\includegraphics[width=3.5in]{../Figures/NCTW_GAS_HF_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_PUF_HF_Center_Avg} \\

	\caption[Wall heat flux along the centerline of the burners]{Wall heat fluxes along the centerline of the burners.  The heat fluxes are averages of replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam are presented.}
	\label{NCTW_HF_Center_Comp}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/NCTW_NG_HF_Offset_Avg} \\
	\includegraphics[width=3.5in]{../Figures/NCTW_GAS_HF_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/NCTW_PUF_HF_Offset_Avg} \\

	\caption[Wall heat flux along the edge of the burners]{Wall heat fluxes along the edge of the burners.  The heat fluxes are averages of replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam are presented.}
	\label{NCTW_HF_Edge_Comp}
\end{figure}
